---
title: "p8105_hw3_nrl2142"
output: github_document
date: "2025-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library (ggridges)
library(patchwork)
library(p8105.datasets)
```
## Problem 1
```{r}
data("instacart")
instacart

instacart|>
  count(aisle_id, sort = TRUE)
```

There are `r n_distinct(instacart$aisle_id)` number of aisles in the instacart dataset. The aisles with the most items ordered from are aisles 83, 24, 123, 120, 21, 115, 84, 107, 91 and 112. 

Making a plot
```{r}
morethantenk_df =
instacart |>
  select(aisle_id)|>
  count(aisle_id, name = "n" ) |>
  filter (n>10000)|>
  arrange(desc(n))

morethantenk_df

ggplot(morethantenk_df, aes(x=n, y=factor(aisle_id))) +
  geom_col() +
  labs (x = "Count", 
        y = "Aisle ID",
        title = "Aisles with more than 10,000 Items")
```

Making a table
```{r}
targetaisles_df =
  instacart |>
  select (aisle, product_name)|>
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
   group_by(product_name, aisle) |>
  summarise(item_count =n ())|>
  arrange (aisle, desc(item_count)) |>
  group_by(aisle) |>
  slice_max(product_name, n=3) |>
  drop_na() 

targetaisles_df
  
  
```
Making the 2x7 table 
```{r}

orderhour_df <- instacart |>
  select(product_name, order_hour_of_day, order_dow) |>
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(mean_value = mean(order_hour_of_day, na.rm = TRUE), .groups = "drop") |>
  mutate(order_dow = recode(as.character(order_dow),
                            `0` = "Sunday",
                            `1` = "Monday",
                            `2` = "Tuesday",
                            `3` = "Wednesday",
                            `4` = "Thursday",
                            `5` = "Friday",
                            `6` = "Saturday")) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_value
  )
orderhour_df
```

## Problem 2
```{r}
zorizipcount_df =
  zorizipcount_df = read_csv ("/Users/gorgeous/p8105_hw2_nrl2142/hw2_data/zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",na = c("NA", ".","")) |>
  janitor::clean_names() |>
   pivot_longer (
  cols = x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to ="zori",
    names_prefix ="x") |>
  count(region_name) |>
  summarise(
    zip_116 = sum (n==116),
zip_under_10 = sum(n <10))

zorizipcount_df
```
## Short description (double check if I really don't have any zipcodes less than 10)

```{r}
zorirental_df=
 zorirental_df = read_csv ("/Users/gorgeous/p8105_hw2_nrl2142/hw2_data/zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",na = c("NA", ".","")) |>
  janitor::clean_names() |>
   pivot_longer (
  cols = x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to ="zori",
    names_prefix ="x") |>
  separate(date, into = c("year", "month", "day"), sep = "_") |>
  mutate (year = as.integer(year),
          month = month.name [as.numeric (month)]) |>
  drop_na(zori)|>
  select(zori,county_name,year)|>
  group_by(county_name, year) |>
  summarise(mean_value = mean(zori), na.rm = TRUE)

zorirental_df
  
```
## Short descriptive paragraph 
Overall, rental prices are increasing across all boroughs in New York from the years 2015 to 2024. The borough that experienced the steepest increase in rental prices is Manhatten with a rent increase of 1056.398, followed by the Bronx with an incrwase of 737.301. This could potentially be due to reasons such as increasing gentrification, inflation or the general appeal to bein close proximity to Manhatten due to the availability of employment opportunities. 

Zip code rental plot
```{r}
zipcoderental_df = read_csv ("/Users/gorgeous/p8105_hw2_nrl2142/hw2_data/zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",na = c("NA", ".","")) |>
  janitor::clean_names() |>
   pivot_longer (
  cols = x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to ="zori",
    names_prefix ="x") |>
  separate(date, into = c("year", "month", "day"), sep = "_") |>
  mutate (year = as.integer(year),
          month = month.name [as.numeric (month)]) |>
  filter(year == 2023) |>
  drop_na(zori)|>
  select(zori,year,month,region_name,county_name)|>
  group_by(month,region_name, county_name) |>
  summarise(mean_value = mean(zori), na.rm = TRUE)

ggplot(zipcoderental_df, (aes(x = month, y = mean_value, group = region_name, color = county_name)))+
         geom_line (alpha = 0.6)+
         labs(
           title = "NYC Rental Prices by Zip Code Across Months in 2023",
           x = "Month",
           y = "Average Rental Price",
           color = "Borough")



zipcoderental_df
```
## Short Descriptive Paragraph (check why all the months in y axis are messed up)
```{r}

```


